---
title: "Social Network - Assignment #1"
author: "Aurora Sterpellone - Gür Piren"
format:
  html:
    self-contained: true
    embed-resources: true 
    toc: true
    toc-location: left
    toc-floating: true
    toc-depth: 3  
    number-sections: true
date: "2025-04-22"
editor_options: 
  markdown: 
    wrap: sentence
---

# SOCIAL NETWORK ANALYSIS - ASSIGNMENT #1 

## MARVEL CHARACTERS - PARTNERSHIPS NETWORK

```{r message=FALSE}

library(igraph)
library(stringr)
library(dplyr)
library(ggraph) # ultimately, might not use it. check at the end
library(tidygraph) # ultimately, might not use it. check at the end
library(ggrepel)
library(extrafont)

```

1. Choose a network dataset

Marvel character partnerships, 2018

2. Indicate the number of nodes should be larger than 300 but smaller than 10000.

The number of nodes is 350.

3. Import that data and create a graph using R

## READING THE DATA

```{r}

marvel_nodes <- read.csv("data/nodes.csv", stringsAsFactors = FALSE, quote = "\"", fill = TRUE)

marvel_links <- read.csv("data/edges.csv", stringsAsFactors = FALSE)

```

## DATA DESCRIPTION

Here we should explain what each column and respective values represent. For instance, 'Group' refers to whether the character is a hero (0), villain (1), or neither (2). We can do this laterrr!

## DATA CLEANING

```{r}
# to get proper x and y coordinates

marvel_nodes$pos_x <- as.numeric(str_extract(marvel_nodes$X_pos, "-?\\d+\\.\\d+"))
marvel_nodes$pos_y <- as.numeric(str_extract_all(marvel_nodes$X_pos, "-?\\d+\\.\\d+") %>% sapply(`[`, 2))

# we don't need the X_pos anymore
marvel_nodes$X_pos <- NULL
```

### Renaming columns for better readability

```{r}

marvel_nodes <- marvel_nodes |> 
  rename(id = X..index,
         character_name = id,
         x = pos_x,
         y = pos_y)

marvel_links <- marvel_links |> 
  rename(source = X..source,
         target = target)
```

## PLOT

```{r}
graph_marvel <- graph_from_data_frame(d = marvel_links, vertices = marvel_nodes, directed = FALSE)

plot(graph_marvel, vertex.label = V(graph_marvel)$character_name, vertex.size = 5, edge.color = "gray")

```

Let's work on a cleaner plot, ideally I wanted to plot it with those characters that has the biggest sizes (number of partnerships) only, but I was not able to position the node labels clearly. maybe we can try later.

For the font of use

```{r message=FALSE, warning=FALSE}

font_import(pattern = "Marvel")  # Looks for "Marvel" in filenames
loadfonts()  # Registers the fonts with R
```


```{r warning=FALSE}
V(graph_marvel)$group <- marvel_nodes$group

# defining a color for each group - heroes '0', villains '1', and neither '2'
group_colors <- c("0" = "lightblue", "1" = "tomato", "2" = "forestgreen")  

# Set font globally for title and legend
par(family = "Marvel")

plot(graph_marvel, 
     vertex.label = NA,                      
     vertex.size = 3,                        # node size
     edge.color = "gray15",                  # link line color
     vertex.color = group_colors[as.character(V(graph_marvel)$group)],  # colors by group
     layout = cbind(V(graph_marvel)$x, V(graph_marvel)$y))  # for the coordinates

title("Marvel Partnerships Network", cex.main = 1.2)

# for the legend
legend("topleft", 
       legend = c("Heroes (0)", "Villains (1)", "Neither (2)"), 
       col = c("lightblue", "tomato", "forestgreen"), 
       pch = 16,  
       bty = "y")
```




